---
title: "FirstActivity"
author: "Jennyfer Chala"
date: "2023-08-26"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This R language practice is performed, where you will know aspects of the data to enter, such as flight, punctuality, times and delay. Using the package **nycflights13** and **tidyverse** in R

```{r message=FALSE}
library(nycflights13) 
library(tidyverse)
```

The flight function is loaded into the package **nycflights13** and flight data set with **flights_data**

```{r}
f_d <- nycflights13::flights
```

# Exercises
# 5.2.4 Exercises: Items 1 and 2

To resolve, use the **filter()** function and select the rows **flights_data** with the value of the column **arr_delay** which is greater than or equal to 2. when stored on **filtered_de flights**.

```{r}
myDf1 <- filter(f_d, arr_delay >= 2)
```

A table is drawn by **Knitr** taking into account the tail number of the aircraft and the time delay in landing

```{r}
library(knitr)
kable(f_d[1:10,c(12,9)],caption = "ARRIVE DELAY", align = "c")
```

Now in the 2 exercise, using **filter()** to select rows where the value in the dest column is equal to "HOU". The result is stored in **filtered_hou_flights**.

```{r}
myDf2 <- filter(f_d,dest =="HOU")
```

```{r}
library(knitr)
kable(myDf2[1:10,c(13,14)],caption = "HOUSTON DESTINY", align = "c")
```

# 5.3.1 Exercises: All Items

```{r}
sorted_flights_missing_first <- flights %>%
  arrange(desc(is.na(dep_time)))
```

In this code.

We purchase flights from the dataset and pass them through the operator **% >%**, allowing sequence operations. Where the function **arrange()** sorts the descending rows according to the column in **s. na(dep_time)**. The column is logical and becomes TRUE if **dep_time** (output time) is absent and FALSE if it is not and those with missing values in **dep_time** appear first.

```{r}
library(knitr)
kable(sorted_flights_missing_first[1:10,c(7,12)],caption = "MISSING DATA FIRST", align = "c")
```

```{r}
most_delayed_flights <- flights %>%
  arrange(desc(arr_delay))
```

Flights **dataset** are purchased by passing through the operator **% >%** by **arrange()** we sort the rows according to the column **arr_delay**. which means that flights with longer delays come first.

```{r}
library(knitr)
kable(most_delayed_flights[1:10,c(6,9,12)],caption = "MOST DELAYED FLIGHTS", align = "c")
```

```{r}
fastest_flights_desc <- flights %>%
  mutate(speed = distance / air_time) %>%
  arrange(desc(speed))
```

The data is taken **`flights`**  by passing through the operator **`%>%`**. For the function **`mutate()`** and create a new column called **`speed`**. Calculating the speed and dividing the column **`distance`** between the column **`air_time`**. To give us the speed of each flight.
And then the **`arrange()`** function is used to sort the rows in descending order according to the **`speed`** column. That flights with the highest speed will appear first.

```{r}
library(knitr)
kable(fastest_flights_desc[1:10,c(12,20)],caption = "FASTEST FLIGHTS", align = "c")
```

```{r}
farthest_flights <- flights %>%
  arrange(desc(distance))
```

The data is taken **`flights`**  by passing through the operator **`%>%`**.The **`arrange()`** function is then used by sorting the rows in descending order based on the column **`distance`**. Making flights with longer distances will appear first.

```{r}
library(knitr)
kable(farthest_flights[1:10,c(12,15)],caption = "FARTHEST FLIGHTS", align = "c")
```

```{r}
closest_flights <- flights %>%
  arrange(distance)
```

The data is taken **`flights`**  by passing through the operator **`%>%`**.The function **`arrange()`** is used by sorting the rows in ascending order according to the column **`distance`**. Making flights with shorter distances will appear first.

```{r}
library(knitr)
kable(closest_flights[1:10,c(12,15)],caption = "CLOSEST FLIGHTS", align = "c")
```

# 5.4.1 Exercises: Items 2, 3, and 4

Add the function **`select()`** to include the name of a variable several times, appearing in the resulting output

```{r message=FALSE}
select(flights, dep_time, dep_time)
```

The column **`dep_time`** is included twice for the output and with the function **`any_of()`** select columns from a dataframe based on a character vector of the column names. It works when you have a column name vector and you want only to get those columns that match any of the names in the vector

```{r message=FALSE}
vars <- c("year", "month", "day", "dep_delay", "arr_delay")
select(flights, any_of(vars))
```
```{r message=FALSE}
select(flights, contains("TIME"))
```

### 5.5.2 Exercises: Items 1 and 2

```{r message=FALSE}
flights_modified <- flights %>%
  mutate(
    dep_time_mins = (dep_time %/% 100) * 60 + dep_time %% 100,
    sched_dep_time_mins = (sched_dep_time %/% 100) * 60 + sched_dep_time %% 100)
```

The **`flights`** data set is taken and the **`%>%`** operator is used to process it.

Using the **`mutate()`** function, new columns are added to the data. In this case, two additional columns are being created: **`dep_time_mins`** and **`sched_dep_time_mins`**. For both columns, calculations are performed to convert time values (stored in HHMM format) to minutes after midnight. For this, the **`%/%`** operation is used to obtain the hours and **`%`** to obtain the minutes.

After executing this code snippet, the **`flights_modified`** dataset will contain the original columns, together with the two new columns, **`dep_time_mins`** and **`sched_dep_time_mins`**, representing the output times (programmed and actual) in minutes counted from midnight.

```{r}
library(knitr)
kable(flights_modified[1:10,c(12,4,5,20,21)],caption = "SCHEDULED DEPARTURE TIME", align = "c")
```

```{r message=FALSE}
comparison_result <- flights_modified %>%
  mutate(arr_dep_time_diff = arr_time - dep_time_mins) %>%
  filter(!is.na(air_time) & !is.na(arr_dep_time_diff)) %>%
  select(air_time, arr_dep_time_diff)
```

The data **`flights_modified`** above is acquired and passed through **`%>%`**. Where the function **`mutate()`** is used Creating a new column called **`arr_dep_time_diff`**. Calculating the difference between arrival time (**`arr_time`**) and departure time in minutes from midnight (**`dep_time_mins`**).

Then  **`filter()`** is used to remove rows where values are missing in columns **`air_time`** or **`arr_dep_time_diff`**. and at the end **`select()`** * is used to choose only columns **`air_time`** and **`arr_dep_time_diff`**.

```{r}
library(knitr)
kable(comparison_result[1:10,c(1,2)],caption = "COMPARISION OF ARRIVES AND DEPARTURES", align = "c")
```

### 5.6.7 Exercises: item 1

- Calculate the median delay experienced on arrival by the set of flights. This gives us a central value that reflects the typical delay at the time of arrival.

- Determine the percentage of flights arriving with an advance or delay of 15 minutes, 30 minutes or even 2 hours compared to your scheduled schedule. This helps us understand how various delay scenarios are distributed within the flight group.

- Calculate the average delay before departure for the set of flights. This provides information about the average delay before a flight takes off.

- Calculate the percentage of flights that are on time (without delay on arrival) and compare it with the percentage of flights that experience significant delays of 2 hours or more. This gives us an idea of how often flights are on time compared to those with substantial delays.

- Create a density chart that illustrates how arrival delays are distributed across all flights. This visual representation helps us identify common delay ranges and detect outliers.

**Question: What's More Important - Arrival Delay or Departure Delay?**

This question seeks to determine which of the two factors, either the delay in arrival or the delay in departure, has the greatest impact on the overall flight experience. This depends on various elements at play. A delay in arrival affects passenger schedules, flight connections and ground transportation plans. Both aspects are crucial, delay and experience but their importance may vary according to passengers' priorities and the nature of their travel plans

### 5.7.1 Exercises: item 2

```{r echo=FALSE}
worst_punctuality <- flights %>%
  group_by(tailnum) %>%
  summarize(
    total_flights = n(),
    punctual_flights = sum(arr_delay <= 0, na.rm = TRUE),
    punctuality_percentage = (punctual_flights / total_flights) * 100
  ) %>%
  arrange(punctuality_percentage) %>%
  filter(!is.na(punctuality_percentage))

worst_punctuality
```

In this code:

Data is acquired **`flights`** * and can be used with  **`%>%`**  for their respective operations. Data are collected by aircraft tail number (**`tailnum`**). Subsequent calculations will be performed separately.
Then the function **`summary()`** Calculate the summary statistics of each aircraft group. What goes in the function **`summary()`**:

  1.   **`total_flights`** is calculated with the function **`n()`**, giving the total number of flights for each aircraft.
  2.   **`punctual_flights`** is calculated with the function **`sum()`**. Counting the number of flights on which the arrival delay (**`arr_delay`**) is less than or equal to 0 indicating the arrivals on time or in advance. The code **`na.rm = TRUE`** determines the missing values in the column **`arr_delay`**.
  3.   **`punctuality_percentage`** is defined as the relationship between point flights and total flights, multiplied by 100 to obtain the total or necessary percentage.

After **`summary()`**, use **`arrange()`** sorting aircraft according to their punctuality percentages in ascending order. To use **`filter()`** * and remove rows where **`punctuality_percentage`** is not available (***`NA`**) Where the resulting data group is called **`wor_punctuality`**, containing aircraft tail numbers, the total number of flights, the number of punctual flights and the corresponding percentage of punctuality.
Ending in the sample of data **`worst_punctuality`** where the tail numbers of aircraft are seen with the lowest punctuality percentages.

```{r}
library(knitr)
kable(worst_punctuality[1:10,c(1,2,3,4)],caption = "WORST PUNCTUALITY TOP", align = "c")
```